---
title: ''
author: ''
date: ''
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)
library(mgcv)
library(dplyr)
library(brms)
library(cmdstanr)
# bring in demog data
#devtools::install_github("nwfsc-cb/srkw-status")
library(kwdemog)
library(ggplot2)
library(viridis)
library(patchwork)
fit_models = TRUE
```

```{r loaddata}
source("code/01_get_data.r")
```

### Survival modeling

We constructed an age based survival model, using census data after 1976, and animals born after 1960 (to avoid uncertainty and biases with older ages; Ward et al. 2013). Previous modelling efforts with these data had used a stage-based model but by not including animals with unknown ages we can fit a model using ages rather than stages (Hilborn et al. 2012; Ward et al. 2013). The base model is

$$logit(\phi_{a,y}) = B_{0,sex_{a,y}} + s(age,k=5) + s(year, k=7) + B_{froh} * Froh_{a}$$
where $\phi_{a,y}$ is the probability of survival of animal $a$ in year $y$, $B_{0, sex_{a,y}}$ is a sex-specific intercept, $s(age,k=5)$ is a smooth function that lets mortality vary non-linearly as a function of age, $s(year, k=7)$ represents a smooth function, here implemented as a penalized smooth with 7 knots (to capture wiggliness that may be influenced by density dependence, prey, and other factors).  We assumed the effect of $F_{ROH}$ was linear, and estimated it with the coefficient $B_{F_{ROH}}$. Results from this model are in Table S1.

As a sensitivity, we conducted a second analysis to include animals without $F_{ROH}$ data. Using a similar approach to Ford et al. (2018), we conducted a 2-stage Bayesian analysis that involved first fitting the survival model without the $F_{ROH}$ coefficient to animals that don't have genetic data (n = 72 animals). The posterior distributions of fixed effects from this model were then used as a prior for a second model, fit to animals with genetic data (n = 86 animals; Table S2). Results from these analyses are qualitatively similar (Table S1--S2). 

### Fecundity modeling

We adopted a similar approach with an age--based model with fecundity data. Because of uncertainty and bias in female ages in the 1970s (Ward et al. 2013), animals born before 1960 were not included. The initial base model constructed was of the form

$$logit(\theta_{a,y}) = B_{0} + B_{1}*age + B_{2}*age^{2} + s(year,k=7) + B_{froh} * Froh_{a}$$
where $\theta_{a,y}$ represents the probability of giving birth. Like the survival model, the $s()$ function represents and estimated smooth function, and the remaining fixed effect coefficients allow for a quadratic effect of age. Like with survival, we assumed a linear effect of $F_{ROH}$ in link-space. Results from this model are in Table S3. As a sensitivity analysis, we also repeated the analysis using a 2-stage model with a prior developed from fitting the model to animals without $F_{ROH}$ data (n=9; Table S4). 

### Estimation

Bayesian Estimation for all models was performed using R (R Core Development Team 2021) and using Stan (Stan Development Team 2021) via the brms (Bürkner 2017) and cmdstanr packages (Gabry and Češnovar 2021). Stan implements Markov chain Monte Carlo (MCMC) using the No-U Turn Sampling (NUTS) algorithm (Hoffman and Gelman, 2014; Carpenter et al., 2017). Each model was run with 4 parallel MCMC chains, for 5000 iterations (5000 warmup). Convergence was assessed by monitoring the lack of divergent transitions, trace plots, and R-hat statistics (Vehtari et al. 2021).

### Effect sizes

Because of negative relationships between $F_{ROH}$ and survival rates, we calculated the effects of Froh for all 3 variables (1Mb, 5Mb, 10Mb) across the relative ranges that exist in our dataset. Using the 10Mb $F_{ROH}$ scenario as a case study, we calculated the annual survival probability for a 20--year old female and male killer whales. As these probabilities are generally very high, we also calculated the cumulative probability for killer whales living to 40 years, across each $F_{ROH}$ metric (results are given in Figures S1--S2). 

As an additional metric, we calculated the lifetime reproductive success (females only) as a function of $F_{ROH}$ values. For each level of $F_{ROH}$, we simulated 
populations of 20000 individuals. For each individual, we simulated stochastic mortality (drawing from the posterior distribution randomly for each animal-year). Females were assigned random births in a similar fashion, and not allowed to give birth in successive years (because of the prolonged 18 month gestation of killer whales). Summary statistics were then applied to the distribution of LRS across simulated individuals (mean, quartiles, etc).

### References  

Bürkner, Paul-Christian. 2017. brms: An R Package for Bayesian
Multilevel Models Using Stan. Journal of Statistical Software, 80(1),
1-28. doi:10.18637/jss.v080.i01

Carpenter, B., A. Gelman, M. Hoffman, D. Lee, B. Goodrich, M. Betancourt, M. Brubaker, J. Guo, P. Li, and A. Riddell. 2017. Stan: A probabilistic programming language. Journal of Statistical Software, 76(1):1–32. 

Ford, M.J., K.M. Parsons, E.J. Ward, J.A. Hempelmann, C.K. Emmons, M.B.
Hanson, K.C. Balcomb, and L.K. Park. 2018. Inbreeding in an endangered killer
whale population. Animal Conservation, 21:423-432.

Gabry, J. and Češnovar, R. 2021. cmdstanr: R Interface to 'CmdStan'.
https://mc-stan.org/cmdstanr, https://discourse.mc-stan.org.
  
Hilborn, R., S.P. Cox, F.M.D. Gulland, D.G. Hankin, N.T. Hobbs, D.E. Schindler, and A.W. Trites. 2012. The Effects of Salmon Fisheries on Southern Resident Killer Whales: Final Report of the Independent Science Panel. Prepared with the assistance of D.R. Marmorek and A.W. Hall, ESSA Technologies Ltd., Vancouver, B.C. for National Marine Fisheries Service (Seattle. WA) and Fisheries and Oceans Canada (Vancouver. BC). xv + 61 pp.  

Hoffman, M. D. and A. Gelman. 2014. The No-U-turn sampler: Adaptively setting path lengths in Hamiltonian Monte Carlo. Journal of Machine Learning Research, 15:1593–1623.

R Core Team. 2021. R: A language and environment for statistical
computing. R Foundation for Statistical Computing, Vienna, Austria. URL
https://www.R-project.org/.

Stan Development Team. 2021. Stan Modeling Language Users Guide and Reference Manual, 2.28. https://mc-stan.org

Vehtari, A., A. Gelman, D. Simpson, B. Carpenter, and P.-C. Bürkner. 2021. Rank-normalization, folding, and localization: An improved R-hat for assessing convergence of MCMC. Bayesian Anal. 16(2): 667-718.

Ward, E.J., M.J. Ford, R.G. Kope, J.K.B. Ford, A. Velez-Espino, C.K. Parken, L.
LaVoy, M.B. Hanson, K.C. Balcomb. 2013. Estimating the impacts of Chinook salmon
abundance and prey removal by ocean fishing on Southern Resident killer whale
population dynamics. U.S. Dept. Commer., NOAA Tech. Memo. NMFS NWFSC-123, 71 p.

```{r surv-1stage}
### This block is for the 1-stage survival model
dat = dplyr::filter(demog, includeSurv==1,
                         !is.na(alive), birth>1960,
                    !is.na(Froh_min10Mb))

if(fit_models) {
source("code/02_fit_surv_1stage.R")
}

```

```{r surv-2stage}
### This block is fitting the 2-stage survival model
prior_dat = dplyr::filter(demog, includeSurv==1,
                         !is.na(alive), birth>1960,
                    is.na(Froh_min10Mb), 
                    animal %in% c("J035-neonate","J031-neonate")==FALSE)
dat = dplyr::filter(demog, includeSurv==1,
                          !is.na(alive), birth>1960,
                     !is.na(Froh_min10Mb))
dat$sexF1M2 = as.factor(dat$sexF1M2)
prior_dat$age2 = prior_dat$age^2
dat$age2 = dat$age^2
set.seed(123)
prior_dat$sexF1M2[which(prior_dat$sexF1M2==0)] = sample(1:2,size=length(which(prior_dat$sexF1M2==0)),replace=T)
prior_dat$sexF1M2 = as.factor(prior_dat$sexF1M2)
prior_dat$froh = 0

if(fit_models) {
source("code/03_fit_surv_2stage.r")
}

```
\break  

```{r results="hide", fec1stage}
# this block is fitting the 1-stage fecundity model
if(fit_models) {
source("code/04_fit_fec_1stage.r")
}
```

```{r fec2stage}
# This block is fitting the prior for the 2-stage fecundity model
if(fit_models) {
source("code/05_fit_fec_2stage-prior.r")
}
```

```{r fec-1stage-all}
# This block is fitting all animals -- not including Froh as covariate
if(fit_models) {
source("code/07_fit_fec_all.r")
}
```

```{r results="hide", fec2stagepost}
# this block is fitting the 2nd part of the 2-stage fecundity model
if(fit_models) {
source("code/06_fit_fec_2stage-posterior.r")
}
```

\break

Table S1. Table of posterior means, 95% credible intervals, and probability of negative values for the 1-stage survival model with $F_{ROH}$ as a covariate.   
```{r tables1}

df = data.frame("Model" = c("Froh_1Mb","Froh_5Mb","Froh_10Mb"),
                "Mean" = NA,
                "Lower95" = NA,
                "Upper95" = NA,
                "PrNeg" = NA
                )
col <- 4
draws = readRDS("survival/1stage_surv_2_bernoulli.rds")
df[1,2:4] = c(mean(draws[,col]), quantile(draws[,col],0.025), quantile(draws[,col],0.975))
df$PrNeg[1] = length(which(draws[,col]<0))/nrow(draws)

draws = readRDS("survival/1stage_surv_3_bernoulli.rds")
df[2,2:4] = c(mean(draws[,col]), quantile(draws[,col],0.025), quantile(draws[,col],0.975))
df$PrNeg[2] = length(which(draws[,col]<0))/nrow(draws)

draws = readRDS("survival/1stage_surv_4_bernoulli.rds")
df[3,2:4] = c(mean(draws[,col]), quantile(draws[,col],0.025), quantile(draws[,col],0.975))
df$PrNeg[3] = length(which(draws[,col]<0))/nrow(draws)
knitr::kable(df, digits=3)
saveRDS(df, "Table_1.rds")
```

Table S2. Table of posterior means, 95% credible intervals, and probability of negative values for the 2-stage survival model with $F_{ROH}$ as a covariate.  
```{r tables1a}
col <- 4
df = data.frame("Model" = c("Froh_1Mb","Froh_5Mb","Froh_10Mb"),
                "Mean" = NA,
                "Lower95" = NA,
                "Upper95" = NA,
                "PrNeg" = NA
                )
draws = readRDS("survival/2stage_surv_2_bernoulli.rds")
df[1,2:4] = c(mean(draws[,col]), quantile(draws[,col],0.025), quantile(draws[,col],0.975))
df$PrNeg[1] = length(which(draws[,col]<0))/nrow(draws)

draws = readRDS("survival/2stage_surv_3_bernoulli.rds")
df[2,2:4] = c(mean(draws[,col]), quantile(draws[,col],0.025), quantile(draws[,col],0.975))
df$PrNeg[2] = length(which(draws[,col]<0))/nrow(draws)

draws = readRDS("survival/2stage_surv_4_bernoulli.rds")
df[3,2:4] = c(mean(draws[,col]), quantile(draws[,col],0.025), quantile(draws[,col],0.975))
df$PrNeg[3] = length(which(draws[,col]<0))/nrow(draws)
knitr::kable(df, digits=3)
saveRDS(df, "Table_S10.rds")
```

Table S3. Table of posterior means, 95% credible intervals, and probability of negative values for the 1-stage fecundity model with $F_{ROH}$ as a covariate.  
```{r tables3}

df = data.frame("Model" = c("Froh_1Mb","Froh_5Mb","Froh_10Mb"),
                "Mean" = NA,
                "Lower95" = NA,
                "Upper95" = NA,
                "PrNeg" = NA
                )
col=5
draws = readRDS("fecundity/1stage_fec_2_bernoulli.rds")
df[1,2:4] = c(mean(draws[,col]), quantile(draws[,col],0.025), quantile(draws[,col],0.975))
df$PrNeg[1] = length(which(draws[,col]<0))/nrow(draws)

draws = readRDS("fecundity/1stage_fec_3_bernoulli.rds")
df[2,2:4] = c(mean(draws[,col]), quantile(draws[,col],0.025), quantile(draws[,col],0.975))
df$PrNeg[2] = length(which(draws[,col]<0))/nrow(draws)

draws = readRDS("fecundity/1stage_fec_4_bernoulli.rds")
df[3,2:4] = c(mean(draws[,col]), quantile(draws[,col],0.025), quantile(draws[,col],0.975))
df$PrNeg[3] = length(which(draws[,col]<0))/nrow(draws)
knitr::kable(df, digits=3)
saveRDS(df, "Table_S12.rds")
```

Table S4. Table of posterior means, 95% credible intervals, and probability of negative values for the 2-stage fecundity model with $F_{ROH}$ as a covariate.   
```{r tables4}

df = data.frame("Model" = c("Froh_1Mb","Froh_5Mb","Froh_10Mb"),
                "Mean" = NA,
                "Lower95" = NA,
                "Upper95" = NA,
                "PrNeg" = NA
                )
col=5
draws = readRDS("fecundity/2stage_fec_2_bernoulli.rds")
df[1,2:4] = c(mean(draws[,col]), quantile(draws[,col],0.025), quantile(draws[,col],0.975))
df$PrNeg[1] = length(which(draws[,col]<0))/nrow(draws)

draws = readRDS("fecundity/2stage_fec_3_bernoulli.rds")
df[2,2:4] = c(mean(draws[,col]), quantile(draws[,col],0.025), quantile(draws[,col],0.975))
df$PrNeg[2] = length(which(draws[,col]<0))/nrow(draws)

draws = readRDS("fecundity/2stage_fec_4_bernoulli.rds")
df[3,2:4] = c(mean(draws[,col]), quantile(draws[,col],0.025), quantile(draws[,col],0.975))
df$PrNeg[3] = length(which(draws[,col]<0))/nrow(draws)
knitr::kable(df, digits=3)
saveRDS(df, "Table_S13.rds")
```

\break 

Table S5. Summary of all animals included in these analyses. 'S_prior' and 'S_post' represents animals in each component of the 2-stage survival model, 'F_prior' and 'F_post' represents animals included in the 2-stage fecundity model. Animals included in the 1-stage version of these models are the same as those included in 'S_post' and 'F_post'.  

```{r tables5}
animals = dplyr::select(demog, animal, birth, death, sexF1M2,Froh_min1Mb) %>%
  dplyr::group_by(animal) %>% 
  dplyr::summarise(Froh = Froh_min1Mb[1],
                   birth=birth[1],
                   death=death[1],
                   sexF1M2=sexF1M2[1]) %>%
  dplyr::rename(Sex = sexF1M2)
animals$Froh = as.character(animals$Froh)  
animals$Froh[which(is.na(animals$Froh))] = ""
animals$Froh[which(animals$Froh!="")] = "x"
animals$Sex[which(animals$Sex==1)] = "F"
animals$Sex[which(animals$Sex==2)] = "M"
animals$Sex[which(animals$Sex%in%c("F","M")==FALSE)] = "Unk"

animals$S_prior = ""
animals$S_post = ""
animals$F_prior = ""
animals$F_post = ""

dat = dplyr::filter(demog, includeSurv==1,
                         !is.na(alive), birth>1960,
                    is.na(Froh_min10Mb),
                    animal %in% c("J035-neonate","J031-neonate")==FALSE)
animals$S_prior[which(animals$animal %in% dat$animal)] = "x"

dat = dplyr::filter(demog, includeSurv==1,
                         !is.na(alive), birth>1960,
                    !is.na(Froh_min10Mb))
animals$S_post[which(animals$animal %in% dat$animal)] = "x"

base_pop = dplyr::filter(demog, sexF1M2==1, includeFec==1,
                         is.na(Froh_min10Mb), !is.na(alive), 
                         age >= 10, age <= 42,
                         !is.na(gave_birth), year>1975, birth > 1960)
animals$F_prior[which(animals$animal %in% base_pop$animal)] = "x"

prior_pop = dplyr::filter(demog, sexF1M2==1, includeFec==1,
                          !is.na(Froh_min10Mb), !is.na(alive),
                          age >= 10, age <= 42,
                          !is.na(gave_birth),year>1975, birth > 1960)
animals$F_post[which(animals$animal %in% prior_pop$animal)] = "x"

animals = dplyr::filter(animals,!is.na(animal))
for(i in 1:nrow(animals)) {
  s = substr(animals$animal[i],2,2)
  if(s=="0") {
    s1 = substr(animals$animal[i],1,1)
    s2 = substr(animals$animal[i],3,4)
    animals$animal[i] = paste0(s1,s2)
  }
}
knitr::kable(animals)

```

\break  

```{r predsurv}
if(fit_models==TRUE) {
source("code/08_pred_surv_fec.r")
}
```

\break 

Figure S1. Survival probabilities and fecundity probabilities as a function of age, for models including 10Mb $F_{ROH}$ values as covariates. Solid lines represent estimates, and shaded region represents 95% credible intervals of the estimated posterior distribution from the stage-based model. Shown are estimates using the year 2000 as a reference year, and the mean of the SRKW population,  $F_{ROH}$=0.0323.  

```{r}

# df = readRDS("fecundity_output_4.rds")
# g2 = ggplot(dplyr::filter(df, year==2000,froh==0,age>=10,age<=43), aes(age, est)) + 
#   geom_ribbon(aes(ymin=lo,ymax=hi),fill="#a6cef9",alpha = 0.2) + 
#   geom_line(col = "#a6cef9") + 
#   #geom_line(aes(Froh_raw, med), col = viridis(1), linetype=2) + 
#   xlim(1,60) + 
#   theme_bw() + 
#   theme(strip.background =element_rect(fill="white"))+
#   ylab("Fecundity") + 
#   xlab("Age") + 
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#  panel.background = element_blank(), axis.line = element_line(colour = "black"))
df_out = readRDS("fecundity_output_4.rds")
df_out$row = seq(1,nrow(df_out))
df = readRDS("fecundity_output_draws_4.rds")
df = dplyr::left_join(df,df_out)
df$row = as.factor(df$row)

df = dplyr::group_by(df, year, froh, age) %>% 
  dplyr::mutate(mean_est = mean(pred_invlogit))

g2_input <- dplyr::filter(df, draws<5000, year==2000,froh==0,age>=10,age<=43)
g2 = ggplot(g2_input, aes(age, pred_invlogit,group=draws)) + 
  #geom_ribbon(aes(ymin=lo,ymax=hi),fill="#a6cef9",alpha = 0.2) + 
  geom_line(col = "dark blue",alpha=0.01,size=0.01) + 
  geom_line(aes(age, mean_est), col = "dark blue") + 
  #geom_line(aes(Froh_raw, med), col = viridis(1), linetype=2) + 
  xlim(1,60) + 
  theme_bw() + 
  theme(strip.background =element_rect(fill="white"))+
  ylab("Fecundity") + 
  xlab("Age") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
 panel.background = element_blank(), axis.line = element_line(colour = "black"))

df_out = readRDS("survival_output_4.rds")
names(df_out)[5] = "froh_std"
df = readRDS("survival_output_draws_4.rds")

df = dplyr::left_join(df,df_out,by="row")
df$row = seq(1,nrow(df))
# summarize animals across years and 
df$Sex = "Female"
df$Sex[which(df$sexF1M2==2)]="Male"

female_draws = dplyr::filter(df, year==2000,froh_std==0, Sex=="Female")
male_draws = dplyr::filter(df, year==2000,froh_std==0, Sex=="Male")

df = readRDS("survival_output_4.rds")
names(df)[5] = "froh_std"
# # summarize animals across years and 
df$Sex = "Female"
df$Sex[which(df$sexF1M2==2)]="Male"
females = dplyr::filter(df, year==2000,froh_std==0, Sex=="Female")
males = dplyr::filter(df, year==2000,froh_std==0, Sex=="Male")

female_draws <- dplyr::group_by(female_draws, age) %>% 
  dplyr::mutate(est = mean(pred_invlogit))

g0_input <- dplyr::filter(female_draws,draws<5000)
g0 = ggplot(data=g0_input, aes(age, pred_invlogit, group=draws)) + 
  geom_line(col = "dark blue",alpha=0.01,size=0.01) + 
  geom_line(aes(age, est),col="dark blue") + 
  #geom_line(col = "dark blue") + 
  theme_bw() + 
  ylab("Female survival") + 
  xlab("Age") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
 panel.background = element_blank(), axis.line = element_line(colour = "black"))

male_draws <- dplyr::group_by(male_draws, age) %>% 
  dplyr::mutate(est = mean(pred_invlogit))

g1_input <- dplyr::filter(male_draws,draws<5000)
g1 = ggplot(g1_input, aes(age, pred_invlogit, group=draws)) + 
  geom_line(col = "dark blue",alpha=0.01,size=0.01) + 
  geom_line(aes(age, est),col="dark blue") +
  theme_bw() + 
  ylab("Male survival") + 
  xlab("Age") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
 panel.background = element_blank(), axis.line = element_line(colour = "black"))

save(g0_input, g1_input, g2_input, file="fig_16.Rdata")
# df = readRDS("survival_output_4.rds")
# names(df)[5] = "froh_std"
# # summarize animals across years and 
# df$Sex = "Female"
# df$Sex[which(df$sexF1M2==2)]="Male"
# 
# females = dplyr::filter(df, year==2000,froh_std==0, Sex=="Female")
# males = dplyr::filter(df, year==2000,froh_std==0, Sex=="Male")
# 
# g0 = ggplot(females, aes(age, est)) + 
#   geom_ribbon(aes(ymin=lo,ymax=hi),fill="#a6cef9",alpha = 0.2) + 
#   geom_line(col = "#a6cef9") + 
#   theme_bw() + 
#   ylab("Female survival") + 
#   xlab("Age") +
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#  panel.background = element_blank(), axis.line = element_line(colour = "black"))
# 
# g1 = ggplot(males, aes(age, est)) + 
#   geom_ribbon(aes(ymin=lo,ymax=hi),fill="#a6cef9",alpha = 0.2) + 
#   geom_line(col = "#a6cef9") + 
#   theme_bw() + 
#   ylab("Male survival") + 
#   xlab("Age") +
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#  panel.background = element_blank(), axis.line = element_line(colour = "black"))


pdf("Figure S16.pdf")
g0/g1/g2
dev.off()
```

```{r}
# calculate cumulative survival probabilities
if(fit_models==TRUE) {
source("code/09_pred_cumulative.r")
}
```


Figure S2. Annual survival probabilities for 20-year old whales, and cumulative survival probabilities to age 40 (2000 used as reference year, Froh = 10Mb). The shaded region in each plot represents 95% credible intervals, solid line represents the posterior mean, and dashed line represents the posterior median. All estimates are scaled relative to 10Mb Froh = 0.0.  

```{r fig.height=4, figss2}
means = c(0.2748935,0.09359155,0.03234788)#c(-1.307479,-2.481795,0.03234788)
sds = c(0.05170584,0.04568827,0.02563815)# c(0.1772409,0.4826107,0.02563815)

d = read.csv("data/FswithLHandPed.csv") 
d = dplyr::select(d, 
                  id, Froh_min1Mb,
                  Froh_min5Mb,
                  Froh_min10Mb)

ii=3 # 1-3, for Mb = 1,5,10
df = readRDS(paste0("survival_output_",4,".rds"))
draws = readRDS(paste0("survival_output_draws_20_",4,".rds"))

df = dplyr::left_join(draws, df)
#df = df[,-6]
df$Froh_raw = froh_sds[2] * df$froh + froh_means[2]

#if(ii < 3) df$Froh_raw = exp(df$Froh_raw)
ii = 2
df = dplyr::filter(df, Froh_raw < apply(d[,2:3],2,max)[ii],
                   Froh_raw > apply(d[,2:3],2,min)[ii])

fem_20 = dplyr::filter(df, year==2000,age==20,sexF1M2==1) %>% dplyr::mutate(lo = lo/est[1], hi = hi/est[1],est = est/est[1], 
        Sex = "Female")

mal_20 = dplyr::filter(df, year==2000,age==20,sexF1M2==2) %>% dplyr::mutate(lo = lo/est[1], hi = hi/est[1],est = est/est[1], 
        Sex = "Male")
annual_20 = rbind(fem_20,mal_20)

d$Froh_raw = d$Froh_min10Mb
d$est = 0

g1 = ggplot(dplyr::filter(rbind(mal_20, fem_20),draws<5000), aes(Froh_raw, pred_invlogit, group=draws)) + 
 #geom_ribbon(aes(ymin=lo,ymax=hi),fill="#a6cef9",alpha = 0.2) + 
 geom_line(col = "darkblue",alpha=0.01,size=0.01) +
 theme_bw() + 
 theme(strip.background =element_rect(fill="white"))+
 ylab("Annual survival (age 20)") + 
  facet_wrap(~Sex, nrow=2) + 
  coord_cartesian(ylim=c(0.85,1.0)) +  
 labs(x=expression(paste(italic("F"),""[ROH]))) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
 panel.background = element_blank(), axis.line = element_line(colour = "black"))# + 
  #geom_rug(data=d, aes(x=Froh_raw, y=0.88),sides="b",col="grey50")

# filter values to those observed in the data
cum_male = readRDS("cumulative_survival_males.rds")
cum_male$fac = as.numeric(factor(cum_male$cov,levels=c("1Mb","5Mb","10Mb")))
cum_male$min = apply(d[,2:4],2,min)[cum_male$fac]
cum_male$max = apply(d[,2:4],2,max)[cum_male$fac]
cum_male = dplyr::filter(cum_male, cov=="10Mb")
cum_male$row = seq(1,nrow(cum_male))
df_all_draws_m = readRDS("cumulative_survival_alldraws_males.rds")
df_all_draws_m = dplyr::filter(df_all_draws_m, ii==3)
df_all_draws_m = dplyr::left_join(df_all_draws_m, cum_male)

# filter values to those observed in the data
cum_female = readRDS("cumulative_survival_females.rds")
cum_female$fac = as.numeric(factor(cum_female$cov,levels=c("1Mb","5Mb","10Mb")))
cum_female$min = apply(d[,2:4],2,min)[cum_female$fac]
cum_female$max = apply(d[,2:4],2,max)[cum_female$fac]
cum_female = dplyr::filter(cum_female, cov=="10Mb")
cum_female$row = seq(1,nrow(cum_female))
df_all_draws_f = readRDS("cumulative_survival_alldraws_females.rds")
df_all_draws_f = dplyr::filter(df_all_draws_f, ii==3)
df_all_draws_f = dplyr::left_join(df_all_draws_f, cum_female)


df_all_draws_m = dplyr::mutate(df_all_draws_m, lo = lo/est[1], hi = hi/est[1],est = est/est[1],
                Sex = "Male") %>% 
  dplyr::filter(Froh_raw < apply(d[,2:4],2,max)[3],
                         Froh_raw > apply(d[,2:4],2,min)[3])
df_all_draws_f = dplyr::mutate(df_all_draws_f, lo = lo/est[1], hi = hi/est[1],est = est/est[1],
                Sex = "Female") %>% 
  dplyr::filter(Froh_raw < apply(d[,2:4],2,max)[3],
                         Froh_raw > apply(d[,2:4],2,min)[3])


g2 = ggplot(rbind(df_all_draws_m,df_all_draws_f), aes(Froh_raw, pred_invlogit, group=draw)) + 
  #geom_ribbon(aes(ymin=lo,ymax=hi),fill="#a6cef9",alpha=0.2) +
  geom_line(col = "darkblue",alpha=0.01,size=0.01) +
  theme_bw() + 
  ylab("Cumulative survival (to age 40)") + 
    facet_wrap(~Sex, nrow=2) + 
 labs(x=expression(paste(italic("F"),""[ROH]))) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
 panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(strip.background =element_rect(fill="white"))# + 
  #geom_rug(data=d, aes(x=Froh_raw),sides="b",col="grey50")


pdf("Figure_S2.pdf")
#(g1 / g2) + (g3 / g4)
g1 + g2
dev.off()
```


Figure S3. Annual survival probabilities for 20-year old whales (2000 used as reference year). The shaded region in each plot represents 95% credible intervals, solid line represents the posterior mean, and dashed line represents the posterior median. All estimates are scaled relative to the minimum observed $F_{ROH}$ value.  

```{r fig.height=4, figs1}

means = c(0.2748935,0.09359155,0.03234788)#c(-1.307479,-2.481795,0.03234788)
sds = c(0.05170584,0.04568827,0.02563815)# c(0.1772409,0.4826107,0.02563815)

d = read.csv("data/FswithLHandPed.csv") 
d = dplyr::select(d, 
                  id, Froh_min1Mb,
                  Froh_min5Mb,
                  Froh_min10Mb)
for(ii in c(1,3)) {

#ii=1 # 1-3, for Mb = 1,5,10
df = readRDS(paste0("survival_output_",ii+1,".rds"))
draws = readRDS(paste0("survival_output_draws_20_",ii+1,".rds"))
df = dplyr::left_join(draws, df)
#df = df[,-6]
df$Froh_raw = froh_sds[c(1,0,2)[ii]] * df$froh + froh_means[c(1,0,2)[ii]]
#if(ii < 3) df$Froh_raw = exp(df$Froh_raw)
df = dplyr::filter(df, Froh_raw < apply(d[,2:3],2,max)[c(1,0,2)[ii]],
                   Froh_raw > apply(d[,2:3],2,min)[c(1,0,2)[ii]])
fem_20 = dplyr::filter(df, year==2000,age==20,sexF1M2==1) %>% 
  dplyr::group_by(draws) %>%
  dplyr::mutate(est = pred_invlogit/pred_invlogit[1], Sex = "Female")
mal_20 = dplyr::filter(df, year==2000,age==20,sexF1M2==2) %>% 
  dplyr::group_by(draws) %>%
  dplyr::mutate(est = pred_invlogit/pred_invlogit[1], Sex = "Male")
annual_20 = rbind(fem_20,mal_20)
annual_20$cov = c("1Mb","5Mb","10Mb")[ii]
if(ii==1) {
  all_df = annual_20
} else {
  all_df = rbind(all_df, annual_20)
}
}


all_df$name = paste0(all_df$Sex," ", all_df$cov)
all_df$name = factor(all_df$name,c("Female 1Mb", "Female 10Mb", "Male 1Mb","Male 10Mb"))

# calculate means
all_df = dplyr::group_by(all_df, name, Froh_raw) %>%
  dplyr::mutate(mean_est = mean(pred_invlogit))
saveRDS(dplyr::filter(all_df, draws<=5000),file="fig_s17.rds")
g1 = ggplot(dplyr::filter(all_df, draws<=5000), aes(Froh_raw, pred_invlogit, group=draws)) + 
 #geom_ribbon(aes(ymin=lo,ymax=hi),fill="#a6cef9",alpha = 0.2) + 
 #geom_line(col = "#a6cef9") + 
 geom_line(col = "darkblue",alpha=0.01,size=0.01) +
  geom_line(aes(Froh_raw, mean_est), col="darkblue") +  
 theme_bw() + 
 theme(strip.background =element_rect(fill="white"))+
  coord_cartesian(ylim=c(0.88,1.0)) + 
 ylab("Annual survival (age 20)") + 
  facet_wrap(~name, nrow=2,scale="free_x") + 
 labs(x=expression(paste(italic("F"),""[ROH]))) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
 panel.background = element_blank(), axis.line = element_line(colour = "black")) 
  
pdf("Figure_S17.pdf")
g1
dev.off()
```

Figure S4. Cumulative survival probabilities to age 40 (2000 used as reference year). The shaded region in each plot represents 95% credible intervals, solid line represents the posterior mean, and dashed line represents the posterior median. All estimates are scaled relative to the minimum observed $F_{ROH}$ value.  

```{r}

for(jj in c(1,3)) {
  
# filter values to those observed in the data
cum_male = readRDS("cumulative_survival_males.rds")
cum_male$fac = as.numeric(factor(cum_male$cov,levels=c("1Mb","10Mb")))
#cum_male$min = apply(d[,2:3],2,min)[cum_male$fac]
#cum_male$max = apply(d[,2:3],2,max)[cum_male$fac]
cum_male = dplyr::filter(cum_male, cov==c("1Mb","10Mb")[c(1,0,2)[jj]])
cum_male$row = seq(1,nrow(cum_male))
df_all_draws_m = readRDS("cumulative_survival_alldraws_males.rds")
df_all_draws_m = dplyr::filter(df_all_draws_m, ii==jj)
df_all_draws_m = dplyr::left_join(df_all_draws_m, cum_male)

# filter values to those observed in the data
cum_female = readRDS("cumulative_survival_females.rds")
cum_female$fac = as.numeric(factor(cum_female$cov,levels=c("1Mb","10Mb")))
#cum_female$min = apply(d[,2:4],2,min)[cum_female$fac]
#cum_female$max = apply(d[,2:4],2,max)[cum_female$fac]
cum_female = dplyr::filter(cum_female, cov==c("1Mb","10Mb")[c(1,0,2)[jj]])
cum_female$row = seq(1,nrow(cum_female))
df_all_draws_f = readRDS("cumulative_survival_alldraws_females.rds")
df_all_draws_f = dplyr::filter(df_all_draws_f, ii==jj)
df_all_draws_f = dplyr::left_join(df_all_draws_f, cum_female)

df_all_draws_m = dplyr::filter(df_all_draws_m, Froh_raw < apply(d[,2:3],2,max)[c(1,0,2)[jj]],
                         Froh_raw > apply(d[,2:3],2,min)[c(1,0,2)[jj]]) %>%
  dplyr::group_by(draw) %>%
  dplyr::mutate(est = pred_invlogit/pred_invlogit[1], Sex = "Male")
df_all_draws_m = dplyr::group_by(df_all_draws_m, Froh_raw) %>%
  dplyr::mutate(mean_est = mean(pred_invlogit))

df_all_draws_f = dplyr::filter(df_all_draws_f, Froh_raw < apply(d[,2:3],2,max)[c(1,0,2)[jj]],
                         Froh_raw > apply(d[,2:3],2,min)[c(1,0,2)[jj]]) %>%
  dplyr::group_by(draw) %>% 
  dplyr::mutate(est = pred_invlogit/pred_invlogit[1], Sex = "Female")
df_all_draws_f = dplyr::group_by(df_all_draws_f, Froh_raw) %>%
  dplyr::mutate(mean_est = mean(pred_invlogit))

df_all = rbind(df_all_draws_m, df_all_draws_f)
df_all$Froh_raw <- df_all$Froh_raw * froh_sds[c(1,0,2)[jj]] + froh_means[c(1,0,2)[jj]]
if(jj==1) {
  all_df = df_all
} else {
  all_df = rbind(all_df, df_all)
}
}
df_all = all_df


df_all$name = paste0(df_all$Sex," ", df_all$cov)

df_all$name = factor(df_all$name, levels = c("Female 1Mb", "Female 10Mb","Male 1Mb", "Male 10Mb"))


df1 <- dplyr::filter(df_all, draw<=5000, name %in% c("Female 1Mb", "Male 1Mb"))
df2 <- dplyr::filter(df_all, draw<=5000, name %in% c("Female 10Mb", "Male 10Mb"))
df1 <- dplyr::filter(df1, Froh < max(d[,2]), Froh > min(d[,2]))
df2 <- dplyr::filter(df2, Froh < max(d[,3]), Froh > min(d[,3]))
df1$Froh_raw <- df1$Froh * froh_sds[1] + froh_means[1]
df2$Froh_raw <- df2$Froh * froh_sds[2] + froh_means[2]
df_all <- rbind(df1, df2)

saveRDS(df_all, "fig_s18.rds")
df_all <- readRDS("fig_s18.rds")


g2 = ggplot(df_all, aes(Froh_raw, pred_invlogit, group = draw)) + 
  #geom_ribbon(aes(ymin=lo,ymax=hi),fill="#a6cef9",alpha=0.2) +
  geom_line(col = "darkblue",alpha=0.01,size=0.01) +
  geom_line(aes(Froh_raw,mean_est), col="darkblue") + 
  theme_bw() + 
  coord_cartesian(ylim=c(0,1.1)) + 
  ylab("Cumulative survival (to age 40)") + 
    facet_wrap(~name, nrow=2,scale="free_x") + 
 labs(x=expression(paste(italic("F"),""[ROH]))) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
 panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(strip.background =element_rect(fill="white"))


pdf("Figure_S18.pdf")
g2
dev.off()
#gridExtra::grid.arrange(g1,g2,g3,g4,nrow=2)

```

\break  

Figure S5. Absolute and relative lifetime reproductive success for SRKW females, calculated via the age-based models in our analysis. 
```{r fig.height=4, figs3, eval=FALSE}
if(fit_models==TRUE) {
source("code/10_simulate_lrs.r")
}
df_all = readRDS("sim_output.rds")
df_all$Froh_raw[1:100] = df_all$Froh[1:100] * froh_sds[1] + froh_means[1]
df_all$Froh_raw[101:200] = df_all$Froh[101:200] * froh_sds[2] + froh_means[2]

#df_all$min = apply(d[,2:3],2,min)[as.numeric(df_all$Froh)]
#df_all$max = apply(d[,2:3],2,max)[as.numeric(df_all$Froh)]
#df_all = dplyr::filter(df_all, Froh_raw < max,
#                         Froh_raw > min)

# filter out values for plotting
indx <- which(df_all$Froh > min(d[,2]) & df_all$Froh < max(d[,2]))
indx <- indx[which(indx <= 100)]
indx2 <- which(df_all$Froh > min(d[,3]) & df_all$Froh < max(d[,3]))
indx2 <- indx2[which(indx2 > 100)]
df_all <- df_all[c(indx,indx2),]

df_all$Froh = (factor(df_all$cov,levels=c("1Mb","10Mb")))

df_all2 = df_all
df_all2 = dplyr::group_by(df_all2, Froh) %>%
  dplyr::mutate(lrs = lrs / lrs[1])
df_all$model = "Absolute"
df_all2$model = "Relative"

df <- rbind(df_all,df_all2)

saveRDS(df, "fig_s19.rds")

g1 = ggplot(df, aes(Froh_raw, lrs,group=Froh, col=Froh)) + 
  #geom_ribbon(aes(ymin=lrs_lo,ymax=lrs_hi),fill=viridis(1),alpha = 0.3) + 
  geom_line(size=1.2) + 
  scale_color_viridis(end=0.8,discrete = TRUE) + 
  #geom_line(aes(Froh_raw, med), col = viridis(1), linetype=2) + 
  theme_bw() + 
  theme(strip.background =element_rect(fill="white"))+
  facet_wrap(~model, scale="free_y",nrow=2) + 
  ylab("Lifetime female reproductive success") + 
  labs(x=expression(paste(italic("F"),""[ROH]))) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
 panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(strip.background =element_rect(fill="white"))
pdf("Figure_S19.pdf")
g1
dev.off()
```

